### 2015-06-14 writing/S'15 LSTM/intro.tex
Abstract and intro revisions

Neural network models that encode sentences as vectors have been successfully used in a wide array of NLP tasks, including translation \cite{sutskever2014sequence}, parsing \cite{dyer2015transition}, and sentiment analysis \cite{tai2015improved}. Many of the most successful of these models are based on either tree models or sequence models. Sequence models based on recurrent neural networks build representations incrementally from left to right \cite{elman1990finding,sutskever2014sequence}, while tree models based on \ii{recursive} neural networks \cite{goller1996learning,socher2011semi} build representations incrementally according the hierarchical sturcture of linguistic phrases. Tree models are presented as the more principled choice, since they align with standard linguistic assumptions about constituent structure and the compositional derivation of complex meanings.

While both classes of model can perform objectively well on many tasks, and both are under active development, tree models have not shown the kinds of dramatic performance improvements over sequence models that their billing would lead one to expect: head-to-head comparisons with sequence models show either modest improvements \cite{tai2015improved} or none at all \cite{li2015tree}. We propose a possible explanation for these results: standard sequence models can learn to implicitly exploit recursive syntactic structure in generating sentence representations, thereby demonstrating the behavior that tree-structured models are explicitly designed around. This would require both that sequence models are able to identify syntactic structure in natural language---which we believe is plausible, on the basis of other recent research---and that they are able to use that structure to guide the way in which they interpret sentences. In this paper, we  evaluate this second ability, focusing on cases where the relevant syntactic structure is clearly indicated in the data.

We compare standard tree and sequence models on their handling of recursive structure by training the models on sentences whose length and recursion depth are limited, and then testing them on longer and more complex sentences, such that only models that exploit the recursive structure will be able to generalize in a way that yields correct interpretations for these test sentences. Our methods are based on those of \newcite{Bowman:Potts:Manning:2014}, who describe an experiment and corresponding artificial dataset which tests this ability in two tree models. We propose a way to adapt than experiment to sequence models by decorating the statements with an explicit bracketing, and use the experiment to compare an LSTM sequence model with the three tree models, with a focus on what data each model needs in order to learn the needed generalizations.

As in Bowman et al., we find that standard tree neural networks are able to make the necessary generalizations, with their performance decaying gradually as the structures in the test set grow in size. We also find that extending the training set to include larger structures mitigates this decay. In addition, we find that a sequence model centered on a single-layer LSTM is also able to generalize to unseen large structures, but that it only does this when trained on a larger and more complex training set than is needed by the tree models.

\paragraph{Related work}
Recent successes on parsing by \newcite{vinyals2014grammar} and \newcite{dyer2015transition} provide evidence that sequence models can learn to recognize syntactic structure in natural language, at least when trained on explicitly syntactic tasks. The simplest model presented in Vinyals et al.~uses an LSTM sequence model to encode each sentence as a vector, and then generates a linearized parse (a sequence of brackets and constituent labels) with high accuracy using only the information present in the vector. This shows that the LSTM was able to identify the correct syntactic structures and also hints that it was able to develop a generalizable method for encoding these structures in vectors. The massive size of the training set that was necessary to train that model---250M tokens---leaves open the possibility that the model primarily learned to generate only tree structures that it had already seen, representing those structures as simple hashes rather than structured objects. If the model were able to map unseen sentences to familiar structures using these hashed representations, it could plausibly have achieved the strong results shown in that paper without showing demonstrating an ability to represent recursive structure, at least in a way that would generalize to substantially new structures.

>>>>

Neural network models that encode sentences as vectors have been successfully used in a wide array of NLP tasks, including translation \cite{sutskever2014sequence}, parsing \cite{dyer2015transition}, and sentiment analysis \cite{tai2015improved}. Many of the most successful of these models are based on either tree or sequence architectures. Sequence models based on recurrent neural networks build representations incrementally from left to right \cite{elman1990finding,sutskever2014sequence}, while tree models based on \ii{recursive} neural networks \cite{goller1996learning,socher2011semi} build representations incrementally according the hierarchical structure of linguistic phrases. Tree models are often presented as the more principled choice, since they align with standard linguistic assumptions about constituent structure and the compositional derivation of complex meanings.

While both classes of model perform %objectively 
well on many tasks, and both are under active development, tree models have not shown the kinds of dramatic performance improvements over sequence models that their billing would lead one to expect: head-to-head comparisons with sequence models show either modest improvements \cite{tai2015improved} or none at all \cite{li2015tree}. We propose a possible explanation for these results: standard sequence models can learn to implicitly exploit recursive syntactic structure in generating sentence representations, thereby demonstrating the behavior that tree-structured models are explicitly designed around. If correct, this means that
%This would require both that 
sequence models are able to identify syntactic structure in natural language---which we believe is plausible, on the basis of other recent research---and that they are able to use that structure to guide interpretation. %the way in which they interpret sentences. 
In this paper, we  evaluate this second ability, focusing on cases where the relevant syntactic structure is clearly indicated in the data.

We compare standard tree and sequence models on their handling of recursive structure by training the models on sentences whose length and recursion depth are limited, and then testing them on longer and more complex sentences, such that only models that exploit the recursive structure will be able to generalize in a way that yields correct interpretations for these test sentences. Our methods are based on those of \newcite{Bowman:Potts:Manning:2014}, who describe an experiment and corresponding artificial dataset which tests this ability in two tree models. We adapt that experiment to sequence models by decorating the statements with an explicit bracketing, and we use this design to compare an LSTM sequence model with three tree models, with a focus on what data each model needs in order to learn the needed generalizations.

As in Bowman et al., we find that standard tree neural networks are able to make the necessary generalizations, with their performance decaying gradually as the structures in the test set grow in size. We also find that extending the training set to include larger structures mitigates this decay. In addition, we find that a sequence model centered on a single-layer LSTM is also able to generalize to unseen large structures, but that it does this only when trained on a larger and more complex training set than is needed by the tree models.

%\paragraph{Related work}
%Recent successes on parsing by 
Out results engage with those of \newcite{vinyals2014grammar} and \newcite{dyer2015transition}, who find that sequence models can learn to recognize syntactic structure in natural language, at least when trained on explicitly syntactic tasks. The simplest model presented in Vinyals et al.~uses an LSTM sequence model to encode each sentence as a vector, and then generates a linearized parse (a sequence of brackets and constituent labels) with high accuracy using only the information present in the vector. This shows that the LSTM was able to identify the correct syntactic structures and also hints that it was able to develop a generalizable method for encoding these structures in vectors. However, the massive size of the data set needed to train the model---250M tokens---leaves open the possibility that it primarily learned to generate only tree structures that it had already seen, representing them as simple hashes rather than structured objects. If the model were able to map unseen sentences to familiar structures using these hashed representations, it could plausibly have achieved the strong results shown in that paper without actually representing recursive structure, at least in a way that would generalize to substantially new structures. Our experiments suggest that LSMTs can generalize in the requisite ways, but at the cost of very large training sets of the right type.
